```{r setup, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# Cross-Validation and the Bootstrap

Dans ce laboratoire, nous explorons les techniques de rééchantillonnage abordées dans ce chapitre. Certaines des commandes de ce laboratoire peuvent prendre un certain temps à s'exécuter sur votre ordinateur.

## L'Approche par Ensemble de Validation

Nous explorons l'utilisation de l'approche par ensemble de validation afin d'estimer les taux d'erreur de test résultant de l'ajustement de divers modèles linéaires sur l'ensemble de données Auto.

Avant de commencer, nous utilisons la fonction set.seed() afin de définir une graine pour le générateur de nombres aléatoires de R, de sorte que le lecteur de ce livre obtiendra précisément les mêmes résultats que ceux montrés ci-dessous. Il est généralement une bonne idée de définir une graine aléatoire lors de la réalisation d'une analyse telle que la validation croisée qui contient un élément d'aléatoire, afin que les résultats obtenus puissent être reproduits précisément à un moment ultérieur.

Nous commençons par utiliser la fonction sample() pour diviser l'ensemble des observations en deux moitiés, en sélectionnant un sous-ensemble aléatoire de 196 observations sur les 392 observations originales. Nous faisons référence à ces observations comme l'ensemble d'entraînement.

```{r chunk1}
library(ISLR2)
set.seed(1)
train <- sample(392, 196)
```

(Ici, nous utilisons un raccourci dans la commande sample ; voir ?sample pour plus de détails.)

Nous utilisons ensuite l'option subset dans lm() pour ajuster une régression linéaire en utilisant uniquement les observations correspondant à l'ensemble d'entraînement.

```{r chunk2}
lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = train)
```

Nous utilisons maintenant la fonction predict() pour estimer la réponse pour toutes les 392 observations, et nous utilisons la fonction mean() pour calculer le MSE des 196 observations dans l'ensemble de validation. Notez que l'index '-train' ci-dessous sélectionne uniquement les observations qui ne sont pas dans l'ensemble d'entraînement.

```{r chunk3}
attach(Auto)
mean((mpg - predict(lm.fit, Auto))[-train]^2)
```

Par conséquent, le MSE de test estimé pour l'ajustement de la régression linéaire est de 23.27. Nous pouvons utiliser la fonction poly() pour estimer l'erreur de test pour les régressions quadratiques et cubiques.

```{r chunk4}
lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, 
    subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)
lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, 
    subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)
```

Ces taux d'erreur sont respectivement de 18.72 et 18.79. Si nous choisissons un ensemble d'entraînement différent, alors nous obtiendrons des erreurs quelque peu différentes sur l'ensemble de validation.

```{r chunk5}
set.seed(2)
train <- sample(392, 196)
lm.fit <- lm(mpg ~ horsepower, subset = train)
mean((mpg - predict(lm.fit, Auto))[-train]^2)
lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, 
    subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)
lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, 
    subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)
```

En utilisant cette division des observations en un ensemble d'entraînement et un ensemble de validation, nous trouvons que les taux d'erreur de l'ensemble de validation pour les modèles avec des termes linéaires, quadratiques et cubiques sont respectivement de 25.73, 20.43 et 20.39.

Ces résultats sont cohérents avec nos découvertes précédentes : un modèle qui prédit mpg à l'aide d'une fonction quadratique de horsepower se comporte mieux qu'un modèle qui implique seulement une fonction linéaire de horsepower, et il y a peu de preuves en faveur d'un modèle qui utilise une fonction cubique de horsepower.

## Validation Croisée Leave-One-Out

L'estimation LOOCV peut être automatiquement calculée pour tout modèle linéaire généralisé en utilisant les fonctions glm() et cv.glm(). Dans le laboratoire du Chapitre 4, nous avons utilisé la fonction glm() pour effectuer une régression logistique en passant l'argument family = "binomial". Mais si nous utilisons glm() pour ajuster un modèle sans passer l'argument family, alors il effectue une régression linéaire, tout comme la fonction lm(). Par exemple,

```{r chunk6}
glm.fit <- glm(mpg ~ horsepower, data = Auto)
coef(glm.fit)
```

et

```{r chunk7}
lm.fit <- lm(mpg ~ horsepower, data = Auto)
coef(lm.fit)
```

produisent des modèles de régression linéaire identiques. Dans ce laboratoire, nous effectuerons une régression linéaire en utilisant la fonction glm() plutôt que la fonction lm() car la première peut être utilisée conjointement avec cv.glm(). La fonction cv.glm() fait partie de la bibliothèque boot.

```{r chunk8}
library(boot)
glm.fit <- glm(mpg ~ horsepower, data = Auto)
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta
```

The `cv.glm()` function produces a list with several components.  The two numbers in the `delta` vector contain the cross-validation results. In this case the numbers are identical (up to two decimal places) and correspond to the LOOCV statistic given in (5.1). Below, we discuss a situation in which the two numbers differ. Our cross-validation estimate for the test error is approximately $24.23$.

We can repeat this procedure for increasingly complex polynomial fits.
 To automate the process, we use the  `for()` function to initiate a *for loop* which iteratively fits polynomial regressions for polynomials of order $i=1$ to $i=10$, computes the associated cross-validation error, and stores it in the $i$th element of the vector `cv.error`.
 We begin by initializing the vector. 

```{r chunk9}
cv.error <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}
cv.error
```

As in Figure 5.4, we see a sharp drop in the estimated test MSE between the linear and quadratic fits, but then no clear improvement from using higher-order polynomials.

## $k$-Fold Cross-Validation

The `cv.glm()` function can also be used to implement $k$-fold CV. Below we use $k=10$, a common choice for $k$, on the `Auto` data set.
We once again set a random seed and initialize a vector in which we will store the CV errors corresponding to the polynomial fits of orders one to ten.

```{r chunk10}
set.seed(17)
cv.error.10 <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K = 10)$delta[1]
}
cv.error.10
```

Notice that the computation time is shorter than that of LOOCV.
(In principle, the computation time for LOOCV for a least squares linear model should be faster than for $k$-fold CV, due to the availability
of the formula (5.2) for LOOCV; however, unfortunately the `cv.glm()` function does not make use of this formula.)
We still see little evidence that using cubic or higher-order polynomial terms leads to lower test error than simply using a quadratic fit.

We saw in Section 5.3.2 that the two numbers associated with `delta` are essentially the same when LOOCV is performed.
When we instead perform $k$-fold CV, then the two numbers associated with `delta` differ slightly. The first is the standard $k$-fold CV estimate,
as in (5.3). The second is a bias-corrected version. On this data set, the two estimates are very similar to each other.

## The Bootstrap


We illustrate the use of the bootstrap in the simple example of Section 5.2, as well as on an example involving estimating the
accuracy of the linear regression model on the `Auto` data set.

### Estimating the Accuracy of a Statistic of Interest

One of the great advantages of the bootstrap approach is that it can be
applied in almost all situations. No complicated mathematical calculations
are required. Performing a bootstrap analysis in `R` entails only two
steps. First, we must create a function that computes the statistic of
interest. Second, we use the `boot()` function, which is part of the `boot` library, to perform the bootstrap by repeatedly
sampling observations from the data set with replacement.

The `Portfolio` data set in the `ISLR2` package is simulated data of $100$ pairs of returns, generated in the fashion described in Section 5.2.
To illustrate the use of the bootstrap on this data, we must first
create a function, `alpha.fn()`, which takes as input the $(X,Y)$ data
as well as a vector indicating which observations should be used to
estimate $\alpha$. The function then outputs the estimate for $\alpha$
based on the selected observations.

```{r chunk11}
alpha.fn <- function(data, index) {
  X <- data$X[index]
  Y <- data$Y[index]
  (var(Y) - cov(X, Y)) / (var(X) + var(Y) - 2 * cov(X, Y))
}
```

This function *returns*, or outputs, an  estimate for $\alpha$ based on applying (5.7) to the observations indexed by the argument `index`.
For instance, the following command tells `R` to estimate $\alpha$ using
all $100$ observations.

```{r chunk12}
alpha.fn(Portfolio, 1:100)
```

The next command  uses the `sample()` function to randomly select
$100$ observations from the range $1$ to $100$, with replacement. This is equivalent
to constructing a new bootstrap data set and recomputing $\hat{\alpha}$
based on the new data set.

```{r chunk13}
set.seed(7)
alpha.fn(Portfolio, sample(100, 100, replace = T))
```

We can implement a bootstrap analysis by performing this command many times, recording all of
the corresponding estimates for $\alpha$, and computing the resulting
standard deviation.
However, the `boot()` function automates this approach. Below we produce $R=1,000$ bootstrap estimates for $\alpha$.


```{r chunk14}
boot(Portfolio, alpha.fn, R = 1000)
```


The final output shows that using the original data, $\hat{\alpha}=0.5758$,
and that the bootstrap estimate for ${\rm SE}(\hat{\alpha})$ is $0.0897$.

### Estimating the Accuracy of a Linear Regression Model

The bootstrap approach can be used  to assess the
variability of the coefficient estimates and predictions from a statistical learning method. Here we use the bootstrap approach in order to assess the variability of
the estimates for $\beta_0$ and $\beta_1$, the intercept and slope terms for the linear regression model
that uses  `horsepower` to predict `mpg` in the `Auto` data set. We will compare the estimates obtained using the bootstrap to those obtained using the formulas
for ${\rm SE}(\hat{\beta}_0)$ and ${\rm SE}(\hat{\beta}_1)$ described
in Section 3.1.2.

We first create a simple function, `boot.fn()`, which takes in the
`Auto` data set as well as a set of indices for the observations, and
returns the intercept and slope estimates for the linear regression model. We then apply this function
to the full set of $392$ observations in order to compute the estimates of $\beta_0$ and $\beta_1$ on the entire data set using the usual linear regression coefficient estimate
formulas from Chapter 3. Note that we do not need the `{` and `}` at the beginning and end of the function because it is only one line long.

```{r chunk15}
boot.fn <- function(data, index)
  coef(lm(mpg ~ horsepower, data = data, subset = index))
boot.fn(Auto, 1:392)
```

 The `boot.fn()` function can also be used in order to create
 bootstrap estimates for the intercept and slope terms by randomly sampling from among the observations with replacement. Here we give two examples.

```{r chunk16}
set.seed(1)
boot.fn(Auto, sample(392, 392, replace = T))
boot.fn(Auto, sample(392, 392, replace = T))
```

Next, we use the `boot()` function to compute the standard errors of 1,000 bootstrap estimates for the intercept and slope terms.

```{r chunk17}
boot(Auto, boot.fn, 1000)
```

This indicates that the bootstrap estimate for ${\rm SE}(\hat{\beta}_0)$ is $0.84$, and that the bootstrap estimate for ${\rm SE}(\hat{\beta}_1)$ is $0.0073$.
As discussed in Section 3.1.2, standard formulas can be used to compute the standard errors for the regression coefficients in a linear model. These can be obtained using the  `summary()` function.

```{r chunk18}
summary(lm(mpg ~ horsepower, data = Auto))$coef
```


The standard error estimates for $\hat{\beta}_0$ and
$\hat{\beta}_1$ obtained using the formulas from
Section 3.1.2 are $0.717$ for the intercept and $0.0064$
for the slope. Interestingly, these are somewhat different from the
estimates obtained using the bootstrap.  Does this indicate a problem
with the bootstrap? In fact, it suggests the opposite.  Recall that
the standard formulas given in Equation 3.8 on page 66 rely on certain assumptions. For example, they depend
on the unknown parameter $\sigma^2$, the noise variance. We then estimate $\sigma^2$
using the RSS. Now although the formulas for the standard errors do not rely on the linear model
being correct, the estimate for $\sigma^2$ does.
We see in
Figure 3.8 on page 92 that there is a non-linear relationship in
the data, and so the residuals from a linear fit will be inflated, and so will $\hat{\sigma}^2$.
Secondly, the standard formulas assume (somewhat unrealistically) that the $x_i$ are fixed, and all the variability comes from the variation in the errors $\epsilon_i$.
 The bootstrap approach does not rely on any of these assumptions, and so it is
likely giving a more accurate estimate of the standard errors of
$\hat{\beta}_0$ and $\hat{\beta}_1$ than is the `summary()`
function.

Below we compute the bootstrap standard error estimates and the standard
linear regression estimates that result from fitting the quadratic model to the data. Since this model provides a good fit to the data (Figure 3.8), there is now a better correspondence between the bootstrap estimates and the standard estimates of ${\rm SE}(\hat{\beta}_0)$, ${\rm SE}(\hat{\beta}_1)$ and ${\rm SE}(\hat{\beta}_2)$.

```{r chunk19}
boot.fn <- function(data, index)
  coef(
      lm(mpg ~ horsepower + I(horsepower^2), 
        data = data, subset = index)
    )
set.seed(1)
boot(Auto, boot.fn, 1000)
summary(
    lm(mpg ~ horsepower + I(horsepower^2), data = Auto)
  )$coef
```


